### 2024-07-15 notebooks/02-modelling-results.qmd
created new plots with all means, pretty clean

---
title: Modelling results
---

```{r}
#| label: setup
#| include: false

library(here)
source(here("scripts/brms-models.R"))
```

:::: {.content-visible when-format="html"}
::: {.callout-note collapse="true"}
# Packages and setup
```r
{{< include ../scripts/_setup.R >}}
```
:::
::::

# Implicit task figures

```{r}
summary(fit_8)
```

```{r}
#| label: estimated-means-fit-8

emmeans_8 <- 
  emmeans(fit_8, c("aphantasia", "congruence"), epred = TRUE) |> 
  summary(point.est = median) |> 
  as_tibble() |> 
  mutate(across(3:5, ~.*1000)) |>
  rename(median = emmean)
```

```{r}
subj_means_8 <-
  df_i_rt |> 
  group_by(subjectid, aphantasia, congruence) |>
  reframe(rt = median(rt)*1000)
```

Without individual subject lines:

```{r}
#| label: fig-8
#| fig-width: 9
#| fig-height: 6

dw <- .4

subj_means_8 |> 
  ggplot(aes(x = congruence, y = rt, color = aphantasia, fill = aphantasia)) +
  # ggplot(aes(x = aphantasia, y = rt, color = congruence, fill = congruence)) +
  geom_violinhalf(
    trim = TRUE,
    flip = c(1, 2),
    alpha = .5,
    scale = "width",
    position = position_dodge(width = dw),
    show.legend = FALSE
  ) +
  geom_quasirandom(
    color = "white",
    pch = 21,
    width = .05,
    size = 1.8,
    alpha = .6,
    show.legend = FALSE,
    dodge.width = .4
  ) +
  # group median lines
  geom_line(
    data = emmeans_8,
    aes(y = median, group = aphantasia),
    position = position_dodge(width = dw),
    size = 1,
    show.legend = FALSE
  ) +
  geom_errorbar(
    data = emmeans_8,
    aes(y = median, ymin = lower.HPD, ymax = upper.HPD),
    position = position_dodge(width = dw),
    width = 0,
    linewidth = 1,
    color = "black",
    show.legend = FALSE
  ) +
  geom_point2(
    data = emmeans_8,
    aes(y = median),
    position = position_dodge(width = dw),
    size = 3,
    shape = 21,
    stroke = 1,
    color = "black"
  ) +
  labs(x = NULL, fill = NULL) +
  scale_y_continuous(
    name = "Response time (ms)",
    breaks = breaks_pretty()
  ) +
  theme(
    axis.ticks.y = element_line(linewidth = 0.5),
    panel.grid.major.y = element_line(linewidth = 0.5, color = "grey90"),
    panel.grid.minor.y = element_line(linewidth = 0.5, color = "grey95"),
    legend.position = "top"
  )
```

With individual subject lines:

```{r}
#| label: fig-8-with
#| fig-width: 10
#| fig-height: 6

alpha_subj <- .3
nudge <- .11

subj_means_8 |> 
  ggplot(aes(x = congruence, y = rt, color = aphantasia, fill = aphantasia)) +
  # ggplot(aes(x = aphantasia, y = rt, color = congruence, fill = congruence)) +
  geom_violinhalf(
    trim = TRUE,
    flip = c(1, 2),
    alpha = .5,
    scale = "width",
    position = position_dodge(width = dw),
    show.legend = FALSE
  ) +
  geom_quasirandom(
    color = "white",
    pch = 21,
    width = .05,
    size = 1.8,
    alpha = .6,
    show.legend = FALSE,
    dodge.width = .4
  ) +
  # individual subject lines
  geom_line(
    data = subj_means_8 |> filter(aphantasia == "Aphantasia"),
    aes(group = interaction(subjectid,aphantasia)),
    position = position_nudge(nudge),
    alpha = alpha_subj,
    show.legend = FALSE
  ) +
    geom_line(
    data = subj_means_8 |> filter(aphantasia == "Control"),
    aes(group = interaction(subjectid, aphantasia)),
    position = position_nudge(-nudge),
    alpha = alpha_subj,
    show.legend = FALSE
  ) +
  # # group median lines
  # geom_line(
  #   data = emmeans_8,
  #   aes(y = median, group = aphantasia),
  #   position = position_dodge(width = dw),
  #   size = 1,
  #   show.legend = FALSE
  # ) +
  geom_errorbar(
    data = emmeans_8,
    aes(y = median, ymin = lower.HPD, ymax = upper.HPD),
    position = position_dodge(width = dw),
    width = 0,
    linewidth = 1,
    color = "black",
    show.legend = FALSE
  ) +
  geom_point2(
    data = emmeans_8,
    aes(y = median),
    position = position_dodge(width = dw),
    size = 3,
    shape = 21,
    stroke = 1,
    color = "black"
  ) +
  labs(x = NULL, fill = NULL) +
  scale_y_continuous(
    name = "Response time (ms)",
    breaks = breaks_pretty()
  ) +
  theme(
    axis.ticks.y = element_line(linewidth = 0.5),
    panel.grid.major.y = element_line(linewidth = 0.5, color = "grey90"),
    panel.grid.minor.y = element_line(linewidth = 0.5, color = "grey95"),
    legend.position = "top"
  )
```

```{r}
#| label: fig-8-zoom
#| fig-width: 10
#| fig-height: 6

subj_means_8 |> 
  ggplot(aes(x = congruence, y = rt, color = aphantasia, fill = aphantasia)) +
  geom_line(
    data = emmeans_8,
    aes(y = median, group = aphantasia),
    position = position_dodge(width = .5),
    size = 1.5,
    show.legend = FALSE
  ) +
  geom_errorbar(
    data = emmeans_8,
    aes(y = median, ymin = lower.HPD, ymax = upper.HPD),
    position = position_dodge(width = .5),
    width = 0,
    linewidth = 1,
    color = "black",
    show.legend = FALSE
  ) +
  geom_point2(
    data = emmeans_8,
    aes(y = median),
    position = position_dodge(width = .5),
    size = 5,
    shape = 21,
    stroke = 1,
    color = "black"
  ) +
  labs(x = NULL, fill = NULL) +
  scale_y_continuous(
    name = "Response time (ms)",
    breaks = breaks_pretty(n = 8)
  ) +
  theme(
    axis.text.y = element_text(size = 10),
    # axis.ticks.y = element_line(linewidth = 0.5),
    panel.grid.major.y = element_line(linewidth = 0.5, color = "grey90"),
    panel.grid.minor.y = element_line(linewidth = 0.5, color = "grey95"),
    legend.position = "top"
  )
```

# Explicit task figures

&nbsp;
&nbsp;
&nbsp;

```{r}
#| label: session-information
#| echo: false

cat("═════════════════════════════════════════════════════════════════════════")
cat("Session information for reproducibility:")
report_system(session = sessionInfo())
cat("Packages used:")
report_packages(session = sessionInfo())
cat("═════════════════════════════════════════════════════════════════════════")
```