### 2024-09-20 notebooks/04-response-letter.qmd
Starting to write the manuscript for Word
- Highlight extension by mcanouil
- _quarto.yml : Adjusted the author list, different for the manuscript and analyses reports
- Transferred Overleaf references
- Added protocol figures
- Integrated Bayesian functions and code to main, need to transfer the bayesian into the modelling report
- Added the old reports temporarily for transferring

---
title: Response to reviewers
subtitle: "Are there unconscious visual images in aphantasia? Development of an implicit priming paradigm"
format: docx
author: "-"
---