### 2024-09-22 notebooks/02-modelling.qmd

---
title: Modelling report
subtitle: "Are there unconscious visual images in aphantasia? Development of an implicit priming paradigm"
# setting for proper numbering of sections
crossref: 
  chapters: true
  custom:
    - kind: float
      key: suppfig
      reference-prefix: Figure S
      space-before-numbering: false
    - kind: float
      key: supptbl
      reference-prefix: Table S
      space-before-numbering: false
supptbl-cap-location: top
---

```{r}
#| label: setup
#| include: false

library(here)
source(here("scripts/00-preprocessing.R"))
```

::: {.content-visible when-format="html"}
::: {.callout-note collapse="true"}
# Packages and setup

``` r
{{< include ../scripts/_setup.R >}}
```
:::
:::

# Questionnaires analyses

## Group differences

### Table

::: {#supptbl-questionnaire-results}
```{r}
#| label: table-questionnaires
#| echo: false

group_effects <- 
  df_questionnaires |>
  pivot_longer(
    cols = c("VVIQ":"SUIS"),
    names_to = "questionnaire",
    values_to = "score"
  ) |> 
  group_by(questionnaire) |> 
  nest() |> 
  rowwise() |> 
  mutate(
    lm = list(lm(score ~ aphantasia + age, data = data)),
    lm_ranked = list(lm(rank(score) ~ aphantasia + age, data = data)),
    ranked_means = list(prepare_means(lm_ranked, questionnaire)),
    table = list(generate_model_summary(lm, lm_ranked, questionnaire))
  )

bind_rows(group_effects$table) |> display()
```
:::

### Plot

```{r}
#| label: questionnaire-plot-setup
#| echo: false

# dodge width for all the geoms
dw <- 1
# axis text size
txt <- 14
# legend text size
txt_legend <- 12

norm_means <- 
  df_q_norm |> 
  group_by(aphantasia) |>
  pivot_longer(
    cols = c(VVIQ:SUIS),
    names_to = "Questionnaire",
    values_to = "score"
  ) |> 
  group_by(Questionnaire) |> 
  nest() |> 
  rowwise() |> 
  mutate(
    lm = list(lm(score ~ aphantasia, data = data)),
    means = list(estimate_means(lm, by = "aphantasia"))
  ) |> 
  unnest(means) |> 
  select(!c(data, lm)) |> 
  ungroup() |> 
  mutate(
    Questionnaire = fct_relevel(
      Questionnaire,
      c("VVIQ", "SUIS", "OSIQ_Object", "OSIQ_Spatial")
    ))
```

::: {#suppfig-questionnaire-plot}
```{r}
#| label: plot-questionnaires
#| fig-width: 10
#| fig-height: 5

df_q_norm |>

  # -------------------------- Long format first
  pivot_longer(
    cols = c(VVIQ:SUIS),
    names_to = "Questionnaire",
    values_to = "Score"
  ) |>
  mutate(
    Questionnaire = fct_relevel(
      Questionnaire, 
      c("VVIQ", "SUIS", "OSIQ_Object", "OSIQ_Spatial")
    )) |>

  # -------------------------- Plotting
  ggplot(aes(
    x = Questionnaire, 
    y = Score, 
    fill = aphantasia, 
    color = aphantasia
    )) +

  # accentuation of the median line at 0.5
  geom_hline(
    yintercept = .5,
    linetype = 1,
    color = "grey30",
    alpha = .2
  ) +

  # --------------------- Scores
  # violin shapes for the distributions and quantiles
  geom_violin(
    position = position_dodge(dw),
    alpha = .1,
    draw_quantiles = c(.25, .5, .75)
  ) +
  # points for means and lines for CIs
  geom_pointrange(
    data = norm_means,
    aes(
      x = Questionnaire,
      y = Mean,
      ymin = CI_low,
      ymax = CI_high,
      color = aphantasia
    ),
    size = .75,
    linewidth = 1,
    position = position_dodge(dw)
  ) +

  # --------------------- Stars and lines
  # stars for significance
  annotate(
    geom  = "text", label = "***", color = "black",
    x = 1, y = .95, size  = 10 
    ) +
  # dotted line that separates the different questionnaires
  geom_vline(
    aes(xintercept = stage(Questionnaire, after_scale = 1.5)),
    linetype = 3
  ) +
  # stars 
  annotate(
    geom  = "text", label = "***", color = "black",
    x = 2, y = .95, size  = 10 
    ) +
  # dotted line
  geom_vline(
    aes(xintercept = stage(Questionnaire, after_scale = 2.5)),
    linetype = 3
  ) +
  # stars
  annotate(
    geom  = "text", label = "***", color = "black",
    x = 3, y = .95, size  = 10 
    ) +
  # dotted line
  geom_vline(
    aes(xintercept = stage(Questionnaire, after_scale = 3.5)),
    linetype = 3
  ) +
  # stars
  annotate(
    geom  = "text", label = "***", color = "black",
    x = 4, y = .95, size  = 10 
    ) +

  # --------------------- Aesthetics
  scale_color_okabeito(name = "Group:  ", labels = c(" Control   ", " Aphantasia")) +
  scale_fill_okabeito(name = "Group:  ", labels  = c(" Control   ", " Aphantasia")) +
  scale_x_discrete(
    name = "",
    labels = c("VVIQ", "SUIS", "OSIQ-Object", "OSIQ-Spatial")
  ) +
  scale_y_continuous(
    name = "Standardized score",
    breaks = breaks_pretty(8)
  ) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_line(),
    axis.text.x  = element_text(size = txt),
    axis.title.y = element_text(size = txt),
    legend.title = element_text(size = txt_legend),
    legend.text  = element_text(size = txt_legend),
    legend.position = "top"
    )

ggsave(here("figures/questionnaires.png"), dpi = 600)
```
:::

## Congruence effects correlations

### Matrices

```{r}
#| label: tables-corr-questionnaires

corr <- 
  df_questionnaires |> 
    select(aphantasia, "Implicit effect":"SUIS") |>
    correlation(method = "spearman", p_adjust = "fdr")

corr
```

::: {#suppfig-corr-mat-questionnaires}
```{r}
#| label: corr-mat-questionnaires
#| fig-width: 10
#| fig-height: 6

corr |> summary() |> plot() + 
  labs(title = "Correlation matrix, whole group (Spearman FDR corrected)")

# ggsave(here("figures/corr-mat-whole.png"), dpi = 600)
```
:::

### Plots

```{r}
#| label: corr-plots-questionnaires-setup
#| fig-width: 10
#| fig-height: 5
#| include: false

note_size = 6

# Ranked versions
p_corr_vviq <-
  df_q_ranked |> 
  plot_correlations("VVIQ", "Implicit effect") +
  labs(x = "VVIQ (ranked)", y = "Implicit effect") +
  annotate(
    geom  = "text", 
    label = TeX(r"($\rho = 0.21, \textit{p} = 0.02$)"), 
    color = "black",
    x = 89, y = .25, 
    size  = note_size 
    )

p_corr_osiq <-
  df_q_ranked |> 
  plot_correlations("OSIQ_Object", "Implicit effect") +
  labs(x = "OSIQ-Object (ranked)", y = "Implicit effect") +
  annotate(
    geom  = "text", 
    label = TeX(r"($\rho = 0.26, \textit{p} = 0.003$)"), 
    color = "black",
    x = 80, y = .25, 
    size  = note_size 
    )

p_corr_suis <-
  df_q_ranked |> 
  plot_correlations("SUIS", "Implicit effect") +
  labs(x = "SUIS (ranked)", y = "Implicit effect") +
  annotate(
    geom  = "text", 
    label = TeX(r"($\rho = 0.27, \textit{p} = 0.002$)"), 
    color = "black",
    x = 81, y = .25, 
    size  = note_size 
    )
```

::: {#suppfig-corr-plots-questionnaires}
```{r}
#| label: corr-plots-questionnaires
#| fig-width: 20
#| fig-height: 15

layout <- "
AAAABBBB
##CCCC##
"

# ((p_corr_vviq + p_corr_osiq) / p_corr_suis) +
(p_corr_vviq + p_corr_osiq + p_corr_suis) +
  plot_layout(design = layout, guides = "collect") &
  theme(
    legend.position = "top", 
    legend.text = element_text(size = 20),
    legend.title = element_blank(),
    axis.title = element_text(size = 16)
    )

# ggsave(here("figures/corr-plots.png"), dpi = 600)
```
:::

# Accuracy analyses

We quickly checked for accuracy differences between groups with a logistic regression model.

```{r}
#| label: accuracy-models

fit_acc_im <- 
  glmer(
    formula = correct_implicit ~ (aphantasia + congruence + color)^3 + (1 | subjectid),
    data = df_i_acc,
    family = binomial(link = "logit"),
  control = glmerControl(optimizer = "bobyqa")
  )

fit_acc_ex <- 
  glmer(
    formula = correct_explicit ~ (aphantasia + congruence + color)^3 + (1 | subjectid),
    data = df_e_acc,
    family = binomial(link = "logit"),
    control = glmerControl(optimizer = "bobyqa")
  )
```

::: {#supptbl-acc-i-summary}
```{r}
#| label: summary-accuracy-implicit
#| echo: false

model_performance(fit_acc_im) |> display()

# model_parameters(fit_acc_im) |> display()

Anova(fit_acc_im)
```

Performance and estimates of the logistic GLMM fitted on accuracy in the implicit task.
:::

::: {#supptbl-acc-i-estimates}
```{r}
#| label: estimates-accuracy-implicit
#| echo: false

estimate_means(fit_acc_im, by = "color", transform = "response") |> display()

estimate_contrasts(fit_acc_im, contrast = "color", transform = "response", p_adjust = "none") |> display()
```

Estimated probability of a correct answer and log-odds contrasts between groups for the accuracy in the implicit task.
:::

::: {#supptbl-acc-e-summary}
```{r}
#| label: summary-accuracy-explicit
#| echo: false

model_performance(fit_acc_ex) |> display()

# model_parameters(fit_acc_ex) |> display()

Anova(fit_acc_ex)
```

Performance and estimates of the logistic GLMM fitted on accuracy in the explicit task.
:::

::: {#supptbl-acc-e-estimates}
```{r}
#| label: estimates-accuracy-explicit
#| echo: false

estimate_means(fit_acc_ex, by = "color", transform = "response") |> display()

estimate_contrasts(fit_acc_ex, contrast = "color", transform = "response", p_adjust = "none") |> display()
```

Estimated probability of a correct answer and log-odds contrasts between groups for the accuracy in the explicit task.
:::

# Response Times analyses

## Rationale

To account for the non-normal, positively skewed distributions of the RTs, we fitted Generalized Linear Mixed Models (GLMMs) with inverse Gaussian distributions. The models were implemented in the `lme4` R package and integrated in `tidymodels` workflows using the package `multilevelmod`. Models with Gamma and Gaussian distributions were also fitted and compared with the AIC and BIC to ensure that we chose the best distribution available.

The models included the ***Group*** (aphantasic, control), ***Congruence*** condition (congruent or incongruent) and ***Color*** condition (color or uncolored) along with all their two and three way interactions as fixed categorical predictors, while ***participants*** have been included as grouping factors (i.e. "random effects"). The random effect structure was chosen by fitting and comparing models with every possible combination of distribution and structure (intercept by participant, congruence or color, slope by participant on congruence and/or color) aiming for the best balance between goodness of fit and parsimony. Complex random-effects structures including various slopes on the factors failed to converge to stable and reliable estimates, hence the optimal models chosen included a single by-participant random intercept.

## Model fitting

The formula of the model fitted is $RT \sim (aphantasia + congruence + color)^3 + (1|subjectid)$. [See the HTML version for the code and details.]{.content-hidden when-format="html"}

::: {.content-visible when-format="html"}
```{r}
#| label: model-fitting
#| code-fold: false

model_formula <- rt ~ (aphantasia + congruence + color)^3 + (1|subjectid)

fit_implicit <- 
  glmer(
    formula = model_formula,
    data = df_i_rt,
    family = inverse.gaussian(link = "identity"),
    control = glmerControl(optimizer = "bobyqa")
  )

fit_explicit <- 
  glmer(
    formula = model_formula,
    data = df_e_rt,
    family = inverse.gaussian(link = "identity"),
    control = glmerControl(optimizer = "bobyqa")
  )
```
:::

## Model diagnostics

The quality checks of the models are displayed in [@suppfig-implicit-checks and @suppfig-explicit-checks]{.content-visible when-format="html"} [@suppfig-implicit-checks-2 and @suppfig-explicit-checks-2]{.content-hidden when-format="html"}.

::: {.content-visible when-format="html"}
::: {.panel-tabset .column-page-inset}
### Implicit task

```{r}
#| label: implicit-checks
#| fig-width: 14
#| fig-height: 12
#| eval: false

# characteristics to check
model_checks <- c("pp_check","homogeneity", "vif", "outliers", "qq", "reqq")

check_model(
  fit_implicit, 
  checks = model_checks, 
  detrend = FALSE, 
  residual_type = "normal"
  ) |> 
  plot()

# ggsave(here("figures/checks-implicit.png"), dpi = 600)
```

![Model assumption checks for the Generalized Linear Mixed Model fit on the RTs in the implicit task.](../figures/checks-implicit.png){#suppfig-implicit-checks}

### Explicit task

```{r}
#| label: explicit-checks
#| fig-width: 14
#| fig-height: 12
#| eval: false

check_model(
  fit_explicit, 
  checks = model_checks, 
  detrend = FALSE, 
  residual_type = "normal"
  ) |> 
  plot()

# ggsave(here("figures/checks-explicit.png"), dpi = 600)
```

![Model assumption checks for the Generalized Linear Mixed Model fit on the RTs in the explicit task.](../figures/checks-explicit.png){#suppfig-explicit-checks}
:::
:::

::: {.content-hidden when-format="html"}
![Model assumption checks for the Generalized Linear Mixed Model fit on the RTs in the implicit task.](../figures/checks-implicit.png){#suppfig-implicit-checks-2}

![Model assumption checks for the Generalized Linear Mixed Model fit on the RTs in the explicit task.](../figures/checks-explicit.png){#suppfig-explicit-checks-2}
:::

## Model summaries

::: {#supptbl-glmm-i-summary}
```{r}
#| label: summary-implicit
#| echo: false

model_performance(fit_implicit) |> display()

# model_parameters(fit_implicit) |> display()

Anova(fit_implicit)
```

Performance and estimates of the GLMM fitted on the implicit task data.
:::

::: {#supptbl-glmm-e-summary}
```{r}
#| label: summary-explicit
#| echo: false

model_performance(fit_explicit) |> display()

# model_parameters(fit_explicit) |> display()

Anova(fit_explicit)
```

Performance and estimates of the GLMM fitted on the explicit task data.
:::

## Estimated means and contrasts

::: {#supptbl-estimates-implicit}
```{r}
#| label: emmeans-implicit
#| echo: false

emmeans_im <-
  emmeans(fit_implicit, c("aphantasia", "congruence")) |>
  summary(point.est = median) |>
  as_tibble() |>
  mutate(across(3:7, ~.*1000)) |>
  rename(median = emmean)

emmeans_im |> 
  rename(Group = 1, Condition = 2, "Median (ms)" = median) |>
  display()

fit_implicit |> 
  estimate_contrasts(
    contrast = c("aphantasia", "congruence"), 
    p_adjust = "none"
    ) |> 
  display()
```

Estimated means for each group in each congruence condition and contrasts between groups and conditions in the implicit task.
:::

::: {#supptbl-estimates-explicit}
```{r}
#| label: emmeans-explicit
#| echo: false

emmeans_ex <-
  emmeans(fit_explicit, c("aphantasia", "congruence")) |>
  summary(point.est = median) |>
  as_tibble() |>
  mutate(across(3:7, ~.*1000)) |>
  rename(median = emmean)

emmeans_ex |> 
  rename(Group = 1, Condition = 2, "Median (ms)" = median) |>
  display()

fit_explicit |> 
  estimate_contrasts(
    contrast = c("aphantasia", "congruence"), 
    p_adjust = "none"
    ) |> 
  display()
```

Estimated means for each group in each congruence condition and contrasts between groups and conditions in the explicit task.
:::

## Visualisations

The figures below are also displayed in the main article. [See the HTML version for the code and details.]{.content-hidden when-format="html"}

```{r}
#| label: fig-im
#| fig-cap: Subject means and model-estimated means per group and condition in the implicit task.
#| fig-width: 18
#| fig-height: 7

subj_means_im <-
  df_i_rt |> 
  group_by(subjectid, aphantasia, congruence) |>
  reframe(rt = mean(rt)*1000)

(
  plot_models_full(subj_means_im, "aphantasia", emmeans_im, y_min = 400, y_max = 1110, size = 4) +
  plot_models_zoomed(subj_means_im, "aphantasia", emmeans_im, size = 4) &
  theme(
    legend.position = "top",
    legend.title = element_blank(),
    legend.text = element_text(size = 20),
    legend.margin = margin(0, 150, 25, 0),
    axis.title.y = element_text(size = 20),
    axis.text.x = element_text(size = 18)
  )
) +
  plot_layout(
    guides = "collect",
    widths = c(1, 1.2)
  )

# ggsave(here("figures/model-implicit.png"), dpi = 600)
```

```{r}
#| label: fig-ex
#| fig-cap: Subject means and model-estimated means per group and condition in the explicit task.
#| fig-width: 18
#| fig-height: 7

subj_means_ex <-
  df_e_rt |> 
  group_by(subjectid, aphantasia, congruence) |>
  reframe(rt = mean(rt)*1000)

(
  plot_models_full(subj_means_ex, "aphantasia", emmeans_ex, y_min = 350, y_max = 1450, size = 4) +
  plot_models_zoomed(subj_means_ex, "aphantasia", emmeans_ex, size = 4) &
  theme(
    legend.position = "top",
    legend.title = element_blank(),
    legend.text = element_text(size = 20),
    legend.margin = margin(0, 150, 25, 0),
    axis.title.y = element_text(size = 20),
    axis.text.x = element_text(size = 18)
  )
) +
  plot_layout(
    guides = "collect",
    widths = c(1, 1.2)
  )

# ggsave(here("figures/model-explicit.png"), dpi = 600)
```

# Finer sub-groups

::: {#supptbl-finer-check}
```{r}
#| label: checking-finer-groups

df_i_acc |> 
  select(subjectid, aphantasia, sub_group) |> 
  distinct() |> 
  group_by(aphantasia, sub_group) |> 
  count() |> 
  arrange(sub_group) |> 
  rename(Aphantasia = 1, "Sub-group" = 2, N = 3) |> 
  display()
```
:::

Interestingly, we have 39 participants that did not score at floor VVIQ but between 17 and 32, a score range that @reederNonvisualSpatialStrategies2024 have proposed to call "hypophantasia". However, our sample comprised only 2 hyperphantasics, i.e. participants scoring above 74 [@zemanPhantasiaPsychologicalSignificance2020]. These two participants will therefore be removed from the sample before performing new analyses on the aphantasic, hypophantasic and control groups.

## Questionnaires analyses

### Table

::: {#supptbl-questionnaire-results}
```{r}
#| label: table-finer-questionnaires
#| echo: false

group_effects_finer <- 
  df_questionnaires |>
  filter(sub_group != "Hyperphantasia") |>
  pivot_longer(
    cols = c("VVIQ":"SUIS"),
    names_to = "questionnaire",
    values_to = "score"
  ) |> 
  group_by(questionnaire) |> 
  nest() |> 
  rowwise() |> 
  mutate(
    lm = list(lm(score ~ sub_group + age, data = data)),
    lm_ranked = list(lm(rank(score) ~ sub_group + age, data = data)),
    means = list(generate_model_means(lm, questionnaire, group = "sub_group")),
    contrasts = list(
      estimate_contrasts(lm, contrast = "sub_group", p_adjust = "none") |> 
        mutate(Questionnaire = questionnaire)
        ),
    anovas = list(Anova(lm_ranked))
    # table = list(generate_model_summary(lm, lm_ranked, questionnaire, group = "sub_group"))
  )

bind_rows(group_effects_finer$means) |> display()
bind_rows(group_effects_finer$contrasts) |> display()
```
:::

### Plot

```{r}
#| label: questionnaire-plot-setup-finer
#| echo: false

# dodge width for all the geoms
dw <- 1
# axis text size
txt <- 14
# legend text size
txt_legend <- 12

norm_means_finer <- 
  df_q_norm |> 
  filter(sub_group != "Hyperphantasia") |>
  group_by(sub_group) |>
  pivot_longer(
    cols = c(VVIQ:SUIS),
    names_to = "Questionnaire",
    values_to = "score"
  ) |> 
  group_by(Questionnaire) |> 
  nest() |> 
  rowwise() |> 
  mutate(
    lm = list(lm(score ~ sub_group, data = data)),
    means = list(estimate_means(lm, by = "sub_group"))
  ) |> 
  unnest(means) |> 
  select(!c(data, lm)) |> 
  ungroup() |> 
  mutate(
    Questionnaire = fct_relevel(
      Questionnaire,
      c("VVIQ", "SUIS", "OSIQ_Object", "OSIQ_Spatial")
    ))

norm_means_finer
```

::: {#suppfig-questionnaire-plot}

```{r}
#| label: plot-questionnaires-finer
#| fig-width: 10
#| fig-height: 5

df_q_norm |> 
  filter(sub_group != "Hyperphantasia") |>

  # --------------------- Long format first
  pivot_longer(
    cols = c(VVIQ:SUIS),
    names_to = "Questionnaire",
    values_to = "Score"
  ) |>
  mutate(
    Questionnaire = fct_relevel(
      Questionnaire, 
      c("VVIQ", "SUIS", "OSIQ_Object", "OSIQ_Spatial")
    )) |>

  # --------------------- Plotting 
  ggplot(aes(
    x = Questionnaire, 
    y = Score, 
    fill = sub_group, 
    color = sub_group
    )) +

  # accentuation of the median line at 0.5
  geom_hline(
    yintercept = .5,
    linetype = 1,
    color = "grey30",
    alpha = .2
  ) +

  # ---------------- Scores
  # violin shapes for the distributions and quantiles
  geom_violin(
    position = position_dodge(dw),
    alpha = .1,
    draw_quantiles = c(.25, .5, .75)
  ) +
  # points for means and lines for CIs
  geom_pointrange(
    data = norm_means_finer,
    aes(
      x = Questionnaire,
      y = Mean,
      ymin = CI_low,
      ymax = CI_high,
      color = sub_group
    ),
    size = .75,
    linewidth = 1,
    position = position_dodge(dw)
  ) +

  # ---------------- Lines
  geom_vline(
    aes(xintercept = stage(Questionnaire, after_scale = 1.5)),
    linetype = 3
  ) +
  geom_vline(
    aes(xintercept = stage(Questionnaire, after_scale = 2.5)),
    linetype = 3
  ) +
  # dotted line
  geom_vline(
    aes(xintercept = stage(Questionnaire, after_scale = 3.5)),
    linetype = 3
  ) +

  # --------------- Significance
  # VVIQ control vs others
  plot_significance_labels(label = "***", size = 7,
    x_star = 0.83, y_star = .93,
    x_line = 0.66, x_line_end = 1,
    y_line = .92
  ) +
  geom_segment(
    x = 0.66, xend = 1.34, 
    y = .91, yend = .91, linewidth = .5, color = "black") +
  # VVIQ hypo vs aph
  plot_significance_labels(label = "***", size = 7,
    x_star = 1.17, y_star = .39,
    x_line = 1, x_line_end = 1.34,
    y_line = .38
  ) +

  # SUIS control vs others
  plot_significance_labels(label = "***", size = 7,
    x_star = 1.83, y_star = .93,
    x_line = 1.66, x_line_end = 2,
    y_line = .92
  ) +
  geom_segment(
    x = 1.66, xend = 2.34, 
    y = .91, yend = .91, linewidth = .5, color = "black") +
  # SUIS hypo vs aph
  plot_significance_labels(label = "*", size = 7,
    x_star = 2.17, y_star = .39,
    x_line = 2, x_line_end = 2.34,
    y_line = .38
  ) +

  # OSIQ-O control vs others
  plot_significance_labels(label = "***", size = 7,
    x_star = 2.83, y_star = .93,
    x_line = 2.66, x_line_end = 3,
    y_line = .92
  ) +
  geom_segment(
    x = 2.66, xend = 3.34, 
    y = .91, yend = .91, linewidth = .5, color = "black") +

  # OSIQ-S control vs aph only
  plot_significance_labels(label = "***", size = 7,
    x_star = 4, y_star = .84,
    x_line = 3.66, x_line_end = 4.34,
    y_line = .83
  ) +
  # OSIQ-S hypo vs aph
  plot_significance_labels(label = "*", size = 7,
    x_star = 4.17, y_star = .78,
    x_line = 4, x_line_end = 4.34,
    y_line = .77
  ) +

  # ----------------------------- Aesthetics 
  scale_color_okabeito(name = "Group:  ", labels = c(" Control   ", " Hypophantasia   ", " Aphantasia")) +
  scale_fill_okabeito(name = "Group:  ", labels  = c(" Control   ", " Hypophantasia   ", " Aphantasia")) +
  scale_x_discrete(
    name = "",
    labels = c("VVIQ", "SUIS", "OSIQ-Object", "OSIQ-Spatial"),
    expand = expansion(mult = c(0, 0))
  ) +
  scale_y_continuous(
    name = "Standardized score",
    breaks = breaks_pretty(8),
    expand = expansion(mult = c(0.05, 0.1))
  ) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_line(),
    axis.text.x  = element_text(size = txt),
    axis.title.y = element_text(size = txt),
    legend.title = element_text(size = txt_legend),
    legend.text  = element_text(size = txt_legend),
    legend.position = "top"
    )

ggsave(here("figures/questionnaires-finer.png"), dpi = 600)
```
:::

## Response Times analysis

## Model fitting

```{r}
#| label: finer-models

model_formula_finer <- rt ~ (sub_group + congruence + color)^3 + (1|subjectid)

fit_finer_i <- 
  glmer(
    formula = model_formula_finer,
    data = df_i_finer,
    family = inverse.gaussian(link = "identity"),
    control = glmerControl(optimizer = "bobyqa")
  )

fit_finer_e <-
  glmer(
    formula = model_formula_finer,
    data = df_e_finer,
    family = inverse.gaussian(link = "identity"),
    control = glmerControl(optimizer = "bobyqa")
  )
```

## Model diagnostics

```{r}
#| label: checks-finer-models
#| fig-width: 14
#| fig-height: 12
#| eval: false

# characteristics to check
model_checks <- c("pp_check","homogeneity", "vif", "outliers", "qq", "reqq")

check_model(
  fit_finer_i, 
  checks = model_checks, 
  detrend = FALSE, 
  residual_type = "normal"
  ) |> 
  plot()

# ggsave(here("figures/checks-finer-implicit.png"), dpi = 600)

check_model(
  fit_finer_e, 
  checks = model_checks, 
  detrend = FALSE, 
  residual_type = "normal"
  ) |> 
  plot()

# ggsave(here("figures/checks-finer-explicit.png"), dpi = 600)
```

![Model assumption checks for the Generalized Linear Mixed Model fit on the RTs in the implicit task with finer sub-groups.](../figures/checks-finer-implicit.png){#suppfig-implicit-checks-finer}

![Model assumption checks for the Generalized Linear Mixed Model fit on the RTs in the explicit task with finer sub-groups.](../figures/checks-finer-explicit.png){#suppfig-explicit-checks-finer}

## Model summaries

::: {#supptbl-im-summary-finer}
```{r}
#| label: summary-finer-im

model_performance(fit_finer_i) |> display()

# model_parameters(fit_finer_i) |> display()

Anova(fit_finer_i)
```
:::

::: {#supptbl-ex-summary-finer}
```{r}
#| label: summary-finer-ex

model_performance(fit_finer_e) |> display()

# model_parameters(fit_finer_e) |> display()

Anova(fit_finer_e)
```
:::

## Estimated means and contrasts

::: {#supptbl-im-estimates-finer}
```{r}
#| label: emmeans-finer-im
#| echo: false

emmeans_finer_im <-
  emmeans(fit_finer_i, c("congruence", "sub_group")) |>
  summary(point.est = median) |>
  as_tibble() |>
  mutate(across(3:7, ~.*1000)) |>
  rename(median = emmean)

emmeans_finer_im |>
  rename(Condition = 1, "Sub-group" = 2, "Median (ms)" = median) |>
  display()

fit_finer_i |>
  estimate_contrasts(
    contrast = c("congruence", "sub_group"),
    p_adjust = "none"
  ) |> 
  display()
```
:::

::: {#supptbl-ex-estimates-finer}
```{r}
#| label: emmeans-finer-ex
#| echo: false

emmeans_finer_ex <-
  emmeans(fit_finer_e, c("congruence", "sub_group")) |>
  summary(point.est = median) |>
  as_tibble() |>
  mutate(across(3:7, ~.*1000)) |>
  rename(median = emmean)

emmeans_finer_ex |>
  rename(Condition = 1, "Sub-group" = 2, "Median (ms)" = median) |>
  display()

fit_finer_e |>
  estimate_contrasts(
    contrast = c("congruence", "sub_group"),
    p_adjust = "none"
  ) |> 
  display()
```
:::

```{r}
#| label: fig-finer-im
#| fig-width: 18
#| fig-height: 7

subj_means_finer_im <-
  df_i_finer |> 
  group_by(subjectid, sub_group, congruence) |>
  reframe(rt = mean(rt)*1000)

(
  plot_models_full(subj_means_finer_im, "sub_group", emmeans_finer_im, size = 4, dw = .6) +
  plot_models_zoomed(subj_means_finer_im, "sub_group", emmeans_finer_im, size = 4) &
  theme(
    legend.position = "top",
    legend.title = element_blank(),
    legend.text = element_text(size = 20),
    legend.margin = margin(0, 150, 25, 0),
    axis.title.y = element_text(size = 20),
    axis.text.x = element_text(size = 18)
  )
) +
  plot_layout(
    guides = "collect",
    widths = c(1, 1.2)
  )

# ggsave(here("figures/model-implicit-finer.png"), dpi = 600)
```

```{r}
#| label: fig-finer-ex
#| fig-width: 18
#| fig-height: 7

subj_means_finer_ex <-
  df_e_finer |> 
  group_by(subjectid, sub_group, congruence) |>
  reframe(rt = mean(rt)*1000)

(
  plot_models_full(subj_means_finer_ex, "sub_group", emmeans_finer_ex, y_min = 350, y_max = 1450, size = 4, dw = .6) +
  plot_models_zoomed(subj_means_finer_ex, "sub_group", emmeans_finer_ex, size = 4) &
  theme(
    legend.position = "top",
    legend.title = element_blank(),
    legend.text = element_text(size = 20),
    legend.margin = margin(0, 150, 25, 0),
    axis.title.y = element_text(size = 20),
    axis.text.x = element_text(size = 18)
  )
) +
  plot_layout(
    guides = "collect",
    widths = c(1, 1.2)
  )

# ggsave(here("figures/model-explicit-finer.png"), dpi = 600)
```

# Bayesian modelling

We used Bayesian hypothesis testing to assess the congruence effect separately for each group. Thus, for each subset of the data (control, aphantasia, hypophantasia, aphantasia without hypo., control without hyperphantasia), we fitted a model with the formula $RT \sim (congruence + color)^3 + (1|subjectid)$ and compared the models with and without the congruence effect using Bayes Factors.

```{r}
#| label: model-formula
#| code-fold: false

formula <- bf(rt ~ congruence * color + (1 | subjectid))
```

We used more sensible priors than the default ones as regularizing priors to avoid overfitting.

```{r}
#| label: priors
#| code-fold: false

# to see the default priors:
# get_prior(formula, data = df_i_rt, family = shifted_lognormal())
# need to define better priors for the intercept, sigma, b, and sd

prior <- c(
  prior(normal(-1, 0.5), class = 'Intercept'),  # around exp(-1) = 0.36 secs
  prior(normal(0.4, 0.3), class = 'sigma'), # SD of individual rts in log-units
  prior(normal(0, 0.3), class = 'b'), # around exp(-1) - exp(-1 + 0.3) = 150 ms 
  prior(normal(0.3, 0.1), class = 'sd')  # some variability between participants
)
```

Bayesian models were fitted using the `brms` package with the `cmdstanr` backend. We used parallel processing to fit the models. We aimed for 40000 post-warmup iterations to have sufficient draws to compute reliable Bayes Factors. Using a computer with 20 cores available for parallel processing, we could fit 20 chains of 3000 iterations with 1000 warmup iterations. Each model fitted in approximately 1 minute. Model fits are saved as `.rds` files for reproducibility and can be found on OSF in the analysis project, under the `data/r-data-structures` folder.

```{r}
#| label: brms-setup

# detecting the number of cores to use
n_cores <- parallel::detectCores() - 4

# defining the number of iterations per chain (+ 1000 warm-up)
n_iter <- ceiling(40000 / n_cores) + 1000

# cmdstanr options
options(cmdstanr_warn_inits = FALSE)
```

```{r}
#| label: model-fit
#| code-summary: Fitting the first model

# Implicit task, aphantasia group ---------------------------------------------
fit_implicit_aph <- 
  brm(
    formula = formula,
    data = df_i_rt |> filter(aphantasia == "Aphantasia"),
    family = shifted_lognormal(),
    prior = prior,
    sample_prior = TRUE,
    chains = n_cores,
    cores  = n_cores,
    iter   = n_iter,
    warmup = 1000,
    refresh = 5,
    backend = "cmdstanr",
    stan_model_args = list(stanc_options = list("O1")),
    save_pars = save_pars(all = TRUE),
    file = here("data/r-data-structures/brms-model-implicit-aph.rds"),
    file_compress = "xz"
  )
```

To save compilation time, we used the `brms` function `update` to use the first compiled model as a template for all the others.

```{r}
#| label: model-updates
#| code-summary: Updating the first model for the other groups

# Implicit task models ----------------------------------------------------

# Control group ---------------------------------------------
fit_implicit_con <-
  # the `custom_update` function can be found in the `scripts/_functions.R` file
  custom_update(fit_implicit_aph,
    data = df_i_rt |> filter(aphantasia == "Control"),
    file = "data/r-data-structures/brms-model-implicit-con.rds"
  )

# Finer aphantasia group -------------------------------------
fit_implicit_finer_aph <- 
  custom_update(fit_implicit_aph,
    data = df_i_finer |> filter(sub_group == "Aphantasia"),
    file = "data/r-data-structures/brms-model-implicit-finer-aph.rds"
  )

# Hypophantasia group -------------------------------------
fit_implicit_hypo <- 
  custom_update(fit_implicit_aph,
    data = df_i_finer |> filter(sub_group == "Hypophantasia"),
    file = "data/r-data-structures/brms-model-implicit-hypo.rds"
  )

# Finer control group -------------------------------------
fit_implicit_finer_con <- 
  custom_update(fit_implicit_aph,
    data = df_i_finer |> filter(sub_group == "Control"),
    file = "data/r-data-structures/brms-model-implicit-finer-con.rds"
  )

# Explicit task models -----------------------------------------------------

# Aphantasia group ---------------------------------------------
fit_explicit_aph <- 
  custom_update(fit_implicit_aph,
    data = df_e_rt |> filter(aphantasia == "Aphantasia"),
    file = "data/r-data-structures/brms-model-explicit-aph.rds"
  )

# Control group ---------------------------------------------
fit_explicit_con <- 
  custom_update(fit_implicit_aph,
    data = df_e_rt |> filter(aphantasia == "Control"),
    file = "data/r-data-structures/brms-model-explicit-con.rds"
  )

# Finer aphantasia group -------------------------------------
fit_explicit_finer_aph <- 
  custom_update(fit_implicit_aph,
    data = df_e_finer |> filter(sub_group == "Aphantasia"),
    file = "data/r-data-structures/brms-model-explicit-finer-aph.rds"
  )

# Hypophantasia group -------------------------------------
fit_explicit_hypo <- 
  custom_update(fit_implicit_aph,
    data = df_e_finer |> filter(sub_group == "Hypophantasia"),
    file = "data/r-data-structures/brms-model-explicit-hypo.rds"
  )

# Finer control group -------------------------------------
fit_explicit_finer_con <- 
  custom_update(fit_implicit_aph,
    data = df_e_finer |> filter(sub_group == "Control"),
    file = "data/r-data-structures/brms-model-explicit-finer-con.rds"
  )
```
We finally compute Bayes Factor in favour of the null model of an absence of congruence effect for each group.

```{r}
#| label: bayes-factors
#| echo: false

cat("Implicit task models ------------------------------------------------\n")

# Aphantasia group
cat(
  "Aphantasia group, evidence against an implicit congruence effect: \n",
  "BF_01 =",
  hypothesis(fit_implicit_aph, "congruence1 = 0")$hypothesis$Evid.Ratio,
  "\n"
  )

# Control group
cat(
  "Control group, evidence against an implicit congruence effect: \n",
  "BF_01 =",
  hypothesis(fit_implicit_con, "congruence1 = 0")$hypothesis$Evid.Ratio,
  "\n"
  )

# Finer aphantasia group
cat(
  "Finer aphantasia group, evidence against an implicit congruence effect: \n",
  "BF_01 =",
  hypothesis(fit_implicit_finer_aph, "congruence1 = 0")$hypothesis$Evid.Ratio,
  "\n"
  )

# Hypophantasia group
cat(
  "Hypophantasia group, evidence against an implicit congruence effect: \n",
  "BF_01 =",
  hypothesis(fit_implicit_hypo, "congruence1 = 0")$hypothesis$Evid.Ratio,
  "\n"
  )

# Finer control group
cat(
  "Finer control group, evidence against an implicit congruence effect: \n",
  "BF_01 =",
  hypothesis(fit_implicit_finer_con, "congruence1 = 0")$hypothesis$Evid.Ratio,
  "\n"
  )

cat("\nExplicit task models ------------------------------------------------\n")

# Aphantasia group
cat(
  "Aphantasia group, evidence against an explicit congruence effect: \n",
  "BF_01 =",
  hypothesis(fit_explicit_aph, "congruence1 = 0")$hypothesis$Evid.Ratio,
  "\n"
  )

# Control group
cat(
  "Control group, evidence against an explicit congruence effect: \n",
  "BF_01 =",
  hypothesis(fit_explicit_con, "congruence1 = 0")$hypothesis$Evid.Ratio,
  "\n"
  )

# Finer aphantasia group
cat(
  "Finer aphantasia group, evidence against an explicit congruence effect: \n",
  "BF_01 =",
  hypothesis(fit_explicit_finer_aph, "congruence1 = 0")$hypothesis$Evid.Ratio,
  "\n"
  )

# Hypophantasia group
cat(
  "Hypophantasia group, evidence against an explicit congruence effect: \n",
  "BF_01 =",
  hypothesis(fit_explicit_hypo, "congruence1 = 0")$hypothesis$Evid.Ratio,
  "\n"
  )

# Finer control group
cat(
  "Finer control group, evidence against an explicit congruence effect: \n",
  "BF_01 =",
  hypothesis(fit_explicit_finer_con, "congruence1 = 0")$hypothesis$Evid.Ratio,
  "\n"
  )
```

::: {.content-visible when-format="html"}

::: {.callout-note collapse="true"}
# Session information for reproducibility

```{r}
#| label: session-information
#| echo: false

cat("═════════════════════════════════════════════════════════════════════════")
report_system(session = sessionInfo())
cat("Packages used:")
report_packages(session = sessionInfo())
cat("═════════════════════════════════════════════════════════════════════════")
```
:::
:::